The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-01-26 article.tex
Filling the gaps

  This CDK version has isolated a new atom typing framework, removing the perception of
  atom types from the algorithms, allowing the perception to be properly unit tested.
  The new code defined the atom type using an atom type ontology based on the XXXXX.

@@

  The new code base defines XXXX atom types for YYYY elements, including ZZZZ radical
  atom types. A full list of atom types is given in Additional File X.

  The code is tested by FFFF unit tests, and has been tested against ChEBI and PubChem. REFS

@@

  \subsection*{Signatures}
  An implementation has been provided of the Signature structure descriptor for molecules 
  \cite{Faulon2003}. These act as a linear notation - like the SMILES format - 
  for the whole molecule as well as for connected substructures rooted at a single atom. The 
  descriptor can also be canonicalized to provide isomorphism-independent representations 
  \cite{Faulon2004}.

@@

  Initiated by Niels Out ...

@@

%  \subsection*{Figure 2 - Sample figure title}
%      Figure legend text.

>>>>

  This CDK version has isolated a new atom typing framework, removing the perception of
  atom types from various algorithms, allowing the perception to be tested separately.
  The new code defines the atom types using a list that specifies for each type
  the element symbol, hybridization, formal charge, number of lone pairs, and
  an enumeration of the bond orders. This list of properties captures the
  information needed for the various algorithms in the CDK. For example,
  hybridization information can be used in certain aromaticity models (see later),
  and the lone pair information is needed for resonance structure calculation
  needed, for example, for Gasteiger \pi-charges.

@@

@@

  \subsection*{Signatures}

  An implementation has been provided of the Signature structure descriptor for molecules 
  \cite{Faulon2003}. These act as a linear notation - like the SMILES format - 
  for the whole molecule as well as for connected substructures rooted at a single atom. The 
  descriptor can also be canonicalized to provide isomorphism-independent representations 
  \cite{Faulon2004}.

@@

  A new rendering API has been introduced to make the rendering code independent
  from Java widget toolkits. The previous code was tightly linked to the Swing
  toolkit, but other tools use different widget toolkits. For example, Bioclipse
  is based on Eclipse which uses the SWT~\cite{spjuth2007bioclipse}. %FIXME: write SWT in full

  A second new design goal was introduced to balance between size restrictions
  of some use cases, such as Java applets, and the rendering functionality. In
  particular, some functionality, even after Modularization, needed considerable
  parts of the CDK library, making creation if a small-sized applet unfeasible.
  Therefore, the rendering API was modularized to allow splitting up rendering
  functionality into modules, with varying CDK dependencies.

  % DESCRIBE THE CORE APIs

@@

\subsection*{Figure 2 - Atom type information specified for a sp3-hybridized
carbon.}
\begin{verbatim}
  <at:AtomType rdf:ID="C.sp3">
    <at:categorizedAs rdf:resource="&cdkat;C.tetrahedral"/>
    <at:hasElement rdf:resource="&elem;C"/>
    <at:hybridization rdf:resource="&at;sp3"/>
    <at:formalCharge>0</at:formalCharge>
    <at:lonePairCount>0</at:lonePairCount>
    <at:formalBondType rdf:resource="&bo;single"/>
    <at:formalBondType rdf:resource="&bo;single"/>
    <at:formalBondType rdf:resource="&bo;single"/>
    <at:formalBondType rdf:resource="&bo;single"/>
  </at:AtomType>
\end{verbatim}
